---
title: "FARS Data Analysis with RpkgsProj"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

This vignette walks you through using the functions provided by `RpkgsProj` to
analyze and visualize FARS data.

```{r setup}
library(RpkgsProj)
```

# Introduction and Context

The US National Highway Traffic Safety Administration's Fatality Analysis
Reporting System (FARS), is a nationwide census providing the American public
yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

Example data can be found 
[here](https://d3c33hcgiwev3.cloudfront.net/_e1adac2a5f05192dc8780f3944feec13_fars_data.zip?Expires=1670976000&Signature=kiMQNN4tlDNQe-pJID3Epe2bMKSvc2bskmq~Poe2jUuBZ8HaexU4rIGVUmFNHOJm45Xwcs6sJIW0Oe2SCvnTh33gP1GbXz4FPJkpoSmTvurr3eMDfbtb0c9caCeeSGj7P4GCDooqaa9eKJALHVtCVMiXpAZ-MMDCvv9F9QP4wko_&Key-Pair-Id=APKAJLTNE6QMUY6HBC5A).

> **Important note**: To avoid the inclusion of large datasets into the package,
the examples displayed below were not evaluated during the building of the
vignette.

# Reading Data

FARS data files can be imported into your R session with the function
`fars_read()`

```{r}
fars_read("accident_2013.csv.bz2")
```

Alternatively, you could use `fars_read_years()` to provide only the year (or
vector of years) as an argument, and the function will look for the
corresponding file.

```{r}
fars_read_years(2013)
```

# Summarizing Data

You can use the function `fars_summarize_years()`, giving it the year(s) as an
argument to summarize the accidents within each month as counts.

```{r}
fars_summarize_years(2013)
```

# Visualizing Data

The package provides the function `fars_map_state()` in which you could provide
a state number and a year to visualize the accidents within that state.

```{r}
fars_map_state(1, 2013)
```

